---
title: "p8105_hw2_rm3609.Rmd"
author: "Runqi Ma"
date: "2018/10/4"
output: html_document
---

#Set up
```{r}
library(tidyverse)
```

#Problem 1

###Read and clean the data; retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. 
###Convert the entry variable from character (YES vs NO) to a logical variable .
```{r}
ny_subway = 
  read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line:entry, vending, ada) %>% 
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))
ny_subway
```

The dataset contains line, station name, station latitude and longitude, routes served, entry, vending, entrance type, and ADA compliance.

I have cleaned the variable names, selected the variables that I want, and changed the entry variable from character to a logical variable.

The resulting dataset is 1868 x 19.

The data is not tidy yet. Some column names contain variable.


###How many distinct stations are there? 
```{r}
dist_both = distinct(ny_subway, line, station_name, .keep_all = TRUE )
dist_both
```

