---
title: "p8105_hw2_rm3609.Rmd"
author: "Runqi Ma"
date: "2018/10/4"
output: html_document
---

#Set up
```{r}
library(tidyverse)
```

#Problem 1

###Read and clean the data; retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. 
###Convert the entry variable from character (YES vs NO) to a logical variable .
```{r}
ny_subway = 
  read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line:entry, vending, ada) %>% 
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))
ny_subway
```

The dataset contains line, station name, station latitude and longitude, routes served, entry, vending, entrance type, and ADA compliance.

I have cleaned the variable names, selected the variables that I want, and changed the entry variable from character to a logical variable.

The resulting dataset is 1868 x 19.

The data is not tidy yet. Some column names contain variable.


###How many distinct stations are there? 
```{r}
dist_both = distinct(ny_subway, line, station_name, .keep_all = TRUE )
dist_both
```
There are 465 distinct stations.

###How many stations are ADA compliant?
```{r}
count(dist_both, ada)
```
84 stations

###What proportion of station entrances / exits without vending allow entrance?
```{r}
filter(ny_subway, vending == "NO" & entry == "TRUE" )
69/1868
```
3.69% of station entrances / exits without vending allow entrance

###Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?
```{r}
ref_sub = gather(ny_subway, key = route_number, value = route_name, route1:route11) %>% 
distinct(line, station_name, .keep_all = TRUE )
count(ref_sub, route_name)
filter(ref_sub, route_name == "A" & ada == TRUE)
```
60 distinct stations serve the A train.
Of the stations that serve the A train, 17 are ADA compliant.


#problem 2
###clean the Mr. Trash Wheel sheet
```{r}
library(readxl)
tra_whe = read_excel ("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", 
                     sheet = "Mr. Trash Wheel", range = "A2:N257" ) %>% 
  janitor::clean_names() %>% 
  rename(weight = weight_tons, volume = volume_cubic_yards) %>% 
  filter(!is.na(dumpster)) %>% 
  mutate(sports_balls = as.integer(sports_balls))

tra_whe
```

###Read and clean precipitation data for 2016 and 2017. For each, omit rows without precipitation data and add a variable year. Next, combine datasets and convert month to a character variable .
```{r}
prec_17 = read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", 
                     sheet = "2017 Precipitation", range = "A2:B14" ) %>% 
  janitor::clean_names() %>% 
  filter(!is.na(total)) %>% 
  mutate(year = 2017)

prec_16 = read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", 
                     sheet = "2016 Precipitation", range = "A2:B14" ) %>% 
  janitor::clean_names() %>% 
  filter(!is.na(total)) %>% 
  mutate(year = 2016)

prec = bind_rows(prec_16, prec_17) %>% 
  mutate(month = month.name[month])

prec
```

###Write a paragraph about these data. Be sure to note the number of observations in both resulting datasets, and give examples of key variables. For available data, what was the total precipitation in 2017? What was the median number of sports balls in a dumpster in 2016?

#Problem 3

###clean the data
```{r}
devtools::install_github("p8105/p8105.datasets")
library(p8105.datasets)
data(brfss_smart2010)

brfss_smart2010 = janitor::clean_names(brfss_smart2010) %>% 
  rename(resp_id = respid, loccation_abbr = locationabbr, location_desc = locationdesc) %>% 
  filter(topic == "Overall Health")  %>% 
  select(year:location_desc, response, data_value)

brfss_smart2010
```

###structure data so that values for Response (“Excellent” to “Poor”) are column names / variables which indicate the proportion of subjects with each response (which are values of Data_value in the original dataset)
###create a new variable showing the proportion of responses that were “Excellent” or “Very Good”
```{r}
brfss_smart2010 = spread(brfss_smart2010, key = response, value = data_value) %>%
  janitor::clean_names() %>%  
  mutate(exc_or_very_good = excellent + very_good)
brfss_smart2010
```


